---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dplyr.extras

<!-- badges: start -->
<!-- badges: end -->

The goal of dplyr.extras is to ...

## Installation

You can install the development version of dplyr.extras from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("vituri/dplyr.extras")
```

## Motivation

`dplyr.extras` intends to:

- make it easy to apply functions conditionally in the middle of a pipe;
- extend `dplyr::filter` and `dplyr::mutate` to vectors;
- generalize functions that "glue" dataframes to an existing one, like `dplyr::add_count`.

## Examples

For the examples, consider this dataframe:

```{r}
library(dplyr.extras)
set.seed(1)
n = 10
df = 
  dplyr::tibble(
    a = runif(n = n)
    , b = c('a', rep('b', 2), rep('c', 3), rep('d', 4))
    )

df
```

### Conditionally apply a function in the middle of a pipe

Suppose we are in a shiny app and there is a filter in a widget; it displays the letters in `df$b` and also the string `all`. If the selected string is `all`, we don't want to filter our `df`. This can be done in a pipe as this:

```{r}
selected_string = 'all' # from the input$selected_string

df %>% 
  {
    if (selected_string == 'all') {
      (.)
    } else {
      (.) %>% filter(b %in% selected_string)
    }
  } # %>% 
#  ... the rest of the pipe
```

With `dplyr.extras`, this can be a bit shorter:

```{r}
df %>% 
  map_if_true(condition = selected_string != 'all', .f = . %>% filter(b %in% selected_string)) # %>% 
  # ...
```

There is also the variant

```r
df %>% 
  map_if_condition(condition, .f_if_true, .f_if_false)
```

where we can apply two different functions, depending on the truthy of `condition`.

As another example: using `dbplyr` to query a table in a database, we can create a function like this

```{r}
get_data_from_my_table = function(con, var1 = NULL, var2 = NULL, var3 = NULL) {
  tbl(con, 'table_name') %>% 
    map_if_true(!is.null(v1), . %>% filter(Column1 %in% var1)) %>% 
    map_if_true(!is.null(v2), . %>% filter(Column2 %in% var2)) %>% 
    map_if_true(!is.null(v3), . %>% filter(Column3 %in% var3)) %>% 
    # ...
    collect()
}
```

to query data with the filters we give. If `var1` is `NULL`, for example, we won't filter the values in Column1 of our table, and so on.

### `add_count` but without count

Sometimes, given a dataframe `df` like above we would like to add columns that somehow aggregate the dataframe. The function `dplyr::add_cout` does this with count:

```{r}
df %>% add_count(b, name = 'count')
```

What if we wanted to summarise the max of `a` for each `b` instead of counting? 

```{r}
df %>% 
  left_join(
    df %>% summarise(max = max(a), .by = b)
  )
```

With `dplyr.extras` we can save some lines:
```{r}
df %>% 
  map_then_left_join(.f = \(df) df %>% summarise(max = max(a), .by = b))
```


### Filtering and modifying vectors

```{r}
library(dplyr.extras)
df = tibble(a = 1:10, b = letters[1:10])
```

It is easy to filter some rows that satisfy a certain condition in a dataframe:

```{r}
df %>%
  filter(a >= 5)
```

Sometimes, however, we only have a vector `v` 
```{r}
v = df$a
v
```

instead of a dataframe. To subset it like above, we would need to do something like

```{r}
v[v >= 5]
```



This is a basic example which shows you how to solve a common problem:

```{r example}
library(dplyr.extras)
## basic example code
```
